---
title: "Coolest Thing I've Learned in R"
author: "Josh Baber"
date: "6/28/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ST558 Third Blog Post  
## Getting Data From APIs

By far the coolest thing I've learned so far with R is how to interact with APIs.  Before this course, I did not even know what an API was, but now I see how useful they can be, especially when used with a statistical programming software like R.  It's super easy and fast to grab just about any custom data or information that we want, and then we can use R to perform statistical analyses with graphs, tables, and summaries on that data.  

## A Quick Example of Using and API

Some APIs require keys, but this [pokeAPI](https://pokeapi.co/) does not.  Using the `fromJSON()` function from the `jsonlite` package, we can easily pull data from a URL that we get by reading the docs of the API.  In this example, I can pull a ton of information about the move "Thunder" and print out its accuracy, power, and type.

```{r}
# Read in jsonlite
library(jsonlite)
# Grab information about the move thunder
thunder <- fromJSON("https://pokeapi.co/api/v2/move/thunder")
# Print out thunder's accuracy, power, and type
c(thunder$accuracy, thunder$power, thunder$type$name)
```

