---
title: "Programming Background"
author: "Josh Baber"
date: "6/8/2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Thoughts on R

  My experience with programming started with R about four years ago in undergrad.  Since then I've used some SAS and some Python and have found those languages much easier to learn than when I first learned R.  In the first R course that I took we learned the basics through a tutorial website called DataCamp.  DataCamp provided plenty of examples for each new topic we learned, and I found it super helpful, but often struggled to complete each part.  I've learned over time that programming is just a completely different way of thinking that I was not used to.  Other people I knew it the course that had experience programming with C, Java, or Python kept saying that R was easy, but I did not believe them.
  Since then, I have become much more familiar with R and ST558 is furthering my awareness of all that R is capable of.  I have come to appreciate R for a number of reasons, it's easy to use, pretty fast, and has RStudio.  RStudio is what makes R so great, it keeps track of objects for you, allows you to work on multiple projects with multiple directories, has built in help and documentation, it autofills functions, variables, and arguments, it can connect to GitHub, it has R Markdown which makes it super easy to create reports, and it has the Tidyverse which makes data manipulation and control much easier and intuitive.  I probably forgot some things, but R just has so much going for it, I understand why it has become an industry standard.
  Compared to SAS, I think in some ways SAS is more rigid when performing analysis but R is more rigid when reading in and manipulating data.  When I was in ST555, which is the NC State SAS course, we had to read in some raw data files that were just a mess.  SAS made it possible as you have a lot of flexibility in specifics of how to read it in.  In R, I feel like it does it for you, same with Python.  I have yet to learn of deeper methods for reading in messy data, however, and I'm sure R has more to offer in that regard.  When performing analysis, R is super easy to use and often there is a function or package with a function that will give you exactly what you need with only a little bit of code.  In SAS, I thought it was strange how much work you would have to do before getting some output you want.  For example, you may need several proc steps to transpose, sort, get the mean, and use it all to find some other value, but in R that would be a very short and simple task, potentially only a couple lines of code or arguments.
  I am currently taking ST590, which is a course that teaches basic Python and applies it to data science.  While I have yet to really get into Python's general purpose side, the data science portions of it are very similar to R.  This, of course, was designed to be when R was created.  Still, it has made learning Python much easier.  I would consider R to be of medium skill to learn, it has a lot of intuitive structures, but sometimes it is easy to get bogged down on the details of it.  For example, in the past I've had to write some really long for loops to get something I want, where I felt it was unnecessary.  Still, I am grateful for R and RStudio and the great resources and packages that are available to get things done.  I feel I am decently comfortable and experienced with R, and yet I feel like it still has much more to offer and I have much more to learn.
  
## Example R Markdown Output

```{r}
plot(iris)
```

